---
title: "Rockies-Project"
author: "Ole Roof"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output:
  html_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r knitr-options, include = FALSE}
knitr::opts_chunk$set(fig.align="center",
                      warning = FALSE,
                      message = FALSE,
                      comment = NA)
```

```{r libs, include=FALSE}
library(tidyverse)
library(Lahman)
library(baseballr)
library(lubridate)
library(scales)
library(gganimate)
library(fuzzyjoin)
library(mgcv)
theme_set(theme_bw())
```

# June 29th game; Los Angeles Dodgers vs. Colorado Rockies
## Rockies Win, 5-3
```{r}
df_col_lad <- scrape_statcast_savant(start_date = ymd("2019-06-29"), 
                                   end_date = ymd("2019-06-29"))

df_col_lad <- df_col_lad %>%
  filter(., home_team == "COL", away_team == "LAD")

score <- matrix(c(1,0,0,1,1,0,0,0,0,3,8,1,0,0,2,0,0,3,0,0,"X",5,7,1),ncol=12,byrow=TRUE)
colnames(score) <- c("1","2","3","4","5","6","7","8","9","R","H","E")
rownames(score) <- c("Los Angeles Dodgers","Colorado Rockies")
score <- as.table(score)
knitr::kable(score)
```

# Box Score

```{r}
url_base <- "http://gd2.mlb.com/components/game/mlb/"
url <- paste0(url_base, 
  "year_2019/month_06/day_29/gid_2019_06_29_lanmlb_colmlb_1/boxscore.xml")
box_score <- baseballr::batter_boxscore(url)
knitr::kable(box_score)
```

# Standings

```{r}
knitr::kable(standings_on_date_bref("2019-06-29", "NL West", from = FALSE))
```

# Strike Zone

```{r}
top_zone <- 3.5
bot_zone <- 1.6
left_zone <- -0.75
right_zone <- 0.75
strike_zone_df <- data.frame(
  x = c(left_zone, left_zone, right_zone, right_zone, left_zone),
  y = c(bot_zone, top_zone, top_zone, bot_zone, bot_zone)
)
```

# How did umpires call balls and strikes?

```{r}
nolan_id <- playerid_lookup(last_name = "Arenado", first_name = "Nolan") %>%
  dplyr::pull(., mlbam_id)

df_nolan_bat <- scrape_statcast_savant(start_date = ymd("2019-06-27"), 
                                       end_date = ymd("2019-06-30"), 
                                       playerid = nolan_id)

df_nolan_bat <- df_nolan_bat %>%
  mutate(swing = if_else(description %in% c("ball", "blocked_ball",
                                            "automatic_ball", "called_strike",
                                            "hit_by_pitch"), 0 , 1))

p <- df_nolan_bat %>%
  filter(swing == "0", type != "X") %>%
  ggplot(aes(., x = plate_x, y = plate_z, col = type))

p + geom_point(alpha = 0.5) +
  scale_color_brewer("call", palette = "Set1") +
  geom_path(data = strike_zone_df,aes(x, y), lwd = 1, color = "black") +
  labs(caption = "Data courtesy of MLBAM",
       x = "horizontal location (ft)",
       y = "vertical location (ft)") +
  xlim(-3.5, 3.5) +
  coord_fixed()
```

# Nolan's Swing Probability

```{r}
swing_gam1 <- gam(swing ~ s(plate_x) + s(plate_z),
                 family = binomial,
                 data = df_nolan_bat)


x <- seq(-1.5, 1.5, length = 50)
z <- seq(0.5, 5, length = 50)
swing_predict_data1 <- tibble(plate_x = c(outer(x, z * 0 + 1)),
                             plate_z = c(outer(x * 0 + 1, z)))


swing_preds1 <- predict(swing_gam1, newdata = swing_predict_data1, type = "response")
swing_predict_data1 <- swing_predict_data1 %>%
  mutate(., swing_prob1 = swing_preds1)

p <- ggplot(data = swing_predict_data1,
            aes(x = plate_x, y = plate_z))

p + geom_tile(aes(fill = swing_prob1)) +
  scale_fill_distiller("", palette = "Spectral",
                       direction = -1,
                       limit = c(0,1)) +
  geom_path(data = strike_zone_df, aes(x, y),
            linetype = 2, color = "navy") +
  coord_fixed() +
  labs(x = "horizontal location (ft)",
       y = "vertical location (ft)",
       caption = "Data courtesy of MLBAM",
       title = "When Does Nolan Arenado Swing?")
```


# Daily Stat Function
## While the baseball savant data has a ton of information, it does not have stats per game or cumulative stats for the season

```{r}
savant_daily <- function(data){
  output <- data %>% 
    filter(!is.na(events), events != "caught_stealing_2b") %>% 
    mutate(
      is_ab = if_else(
        events %in% c("strikeout", "field_out", "single", "double", "force_out", 
                      "triple", "home_run", "double_play", "field_error", "grounded_into_double_play",
                      "strikeout_double_play", "fielders_choice_out"),
        TRUE,
        FALSE
      ),
      is_hit = if_else(
        events %in% c("single", "double", "triple", "home_run"), TRUE, FALSE
      ),
      bases = case_when(
        events == "single" ~ 1,
        events == "double" ~ 2,
        events == "triple" ~ 3,
        events == "home_run" ~ 4,
        TRUE ~ 0
      ),
      event_custom = case_when(
        events %in% c("single", "double", "triple", "home_run") ~ events, 
        str_detect(events, "sac") ~ "sacrifice",
        events %in% c("walk", "hit_by_pitch") ~ NA_character_,
        events == "field_error" ~ "error",
        TRUE ~ "out"
      )
    ) %>% 
    group_by(game_date) %>% 
    summarize(
      pa = length(unique(at_bat_number)),
      ab = sum(is_ab),
      hits = sum(is_hit),
      doubles = sum(events == "double"),
      triples = sum(events == "triples"),
      home_runs = sum(events == "home_run"),
      bb = sum(events == "walk"),
      hbp = sum(events == "hit_by_pitch"),
      so = sum(events %in% c("strikeout", "strikeout_double_play")),
      bases = sum(bases)
    ) %>% 
    arrange(game_date) %>% 
    mutate(
      ba = round(hits/ab, 3),
      obp = round((hits + bb + hbp)/(ab + bb + hbp), 3),
      slg = round(bases/ab, 3),
      ops = obp + slg,
      hits_to_date = cumsum(hits),
      bb_to_date = cumsum(bb),
      hbp_to_date = cumsum(hbp),
      ab_to_date = cumsum(ab),
      bases_to_date = cumsum(bases),
      ba_to_date = round(hits_to_date/ab_to_date, 3),
      obp_to_date = round(
        (hits_to_date + bb_to_date + hbp_to_date)/(ab_to_date + bb_to_date + hbp_to_date), 3
      ),
      slg_to_date = round(bases_to_date/ab_to_date, 3),
      ops_to_date = obp_to_date + slg_to_date
    )
  
  return(output)
}
```

```{r}
charlie_id <- playerid_lookup(last_name = "Blackmon", first_name = "Charlie") %>%
  dplyr::pull(., mlbam_id)
df_charlie_bat <- scrape_statcast_savant(start_date = "2019-03-28", 
                                       end_date = "2019-06-29", 
                                       playerid = charlie_id)
charlie_bsvnt_daily <- savant_daily(df_charlie_bat)

knitr::kable(charlie_bsvnt_daily)
```

# Charlie Blackmon Batting Average

```{r}
charlie_rbi <- charlie_bsvnt_daily %>% 
  ggplot(aes(game_date, ba_to_date)) + 
  geom_step(size = 1.5) + 
  theme_bw() + 
  labs(
    y = "Batting Average",
    title = "Charlie Blackmon's batting average during the 2019 season",
    caption = "Data source: baseballsavant.com\n"
  ) + 
  scale_x_date(date_breaks = "2 weeks", date_labels = "%B %d, %Y") + 
  scale_y_continuous(breaks = seq(0, 0.35, 0.05)) + 
  theme(
    plot.title = element_text(size = 14, face = "bold"),
    plot.caption = element_text(size = 10, face = "bold"),
    axis.title.x = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, face = "bold"),
    axis.text.y = element_text(face = "bold")
  ) + 
  transition_reveal(id = game_date, along = game_date) 


animate(charlie_rbi, width = 800, height = 600)
```


# Charlie Blackmon Spray Chart

```{r}
charlie_spray <- df_charlie_bat %>% 
  filter(events %in% c("single", "double", "triple", "home_run")) %>% 
  mutate(events = factor(events, levels = c("single", "double", "triple", "home_run"),
                         labels = c("Single", "Double", "Triple", "Home Run"))) %>% 
  ggspraychart(fill_value = "events", 
               fill_palette = c("Single"="red", "Double"="blue", 
                                "Triple"="yellow", "Home Run"="green"), point_alpha = 0.5) +
  ggtitle("Charlie Blackmon\n") +
  labs(subtitle = "Spray Chart for 2019\n") + 
  theme(
    legend.title = element_blank(),
    plot.title = element_text(size = 14),
    plot.caption = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )
charlie_spray
```

```{r}
charlie_spray_density <- df_charlie_bat %>% 
  filter(events %in% c("single", "double", "triple", "home_run")) %>% 
  mutate(events = factor(events, levels = c("single", "double", "triple", "home_run"),
                         labels = c("Single", "Double", "Triple", "Home Run"))) %>% 
  ggspraychart(fill_value = "events", 
               fill_palette = c("Single"="red", "Double"="blue", 
                                "Triple"="yellow", "Home Run"="green"), density = TRUE) +
  ggtitle("Charlie Blackmon\n") +
  labs(subtitle = "Spray Chart for 2019\n") + 
  theme(
    legend.title = element_blank(),
    plot.title = element_text(size = 14),
    plot.caption = element_text(size = 10, face = "bold"),
    legend.text = element_text(size = 12, face = "bold")
  )
charlie_spray_density
```

# Clayton Kershaw

```{r}
kersh_id <- playerid_lookup(last_name = "Kershaw", first_name = "Clayton")
df_kersh <- scrape_statcast_savant_pitcher(start_date = ymd("2019-06-29"), 
                                          end_date = ymd("2019-06-29"), 
                                          pitcherid = 477132)
df_kersh_grouped <- df_kersh %>%
  group_by(pitch_type) %>%
  summarize(perc = n()/nrow(df_kersh)) %>%
  arrange(desc(-perc)) %>%
  mutate(pitch_type = factor(pitch_type, pitch_type))

p <- ggplot(df_kersh_grouped %>% filter(pitch_type != "null"), 
            aes(x = pitch_type, y = perc))
p + geom_bar(stat = "identity", fill = "#6e0000") +
  scale_y_continuous(labels = scales::percent) +
  coord_flip() +
  labs(x = "pitch type",
       y = "percentage")
```

```{r}
p <- ggplot(df_kersh %>%
              filter(type != "X"),
            aes(x = plate_x,
                y = plate_z, col = type))
p + geom_point(alpha = 0.4) +
  facet_wrap(. ~ pitch_type, ncol = 2) +
  scale_color_brewer("type", palette = "Set1") +
  geom_path(data = strike_zone_df,aes(x, y), lwd = 1.5, color = "red") +
  labs(title = "Pitches Throw by Clayton Kershaw", 
       caption = "Data courtesy of MLBAM",
       x = "horizontal location (ft)",
       y = "vertical location (ft)") +
  xlim(-3.5, 3.5) +
  coord_fixed()
```

